\input texsis
\paper

\overfullrule=0pt
\def\frac#1#2{#1\over #2}
%
\def\bmatrix#1{\left[ \matrix{#1} \right]}
\def\be{{\beta}}
\def\toone{{t+1}}

%\def\frac#1#2{#1\over #2}
%\magnification=1200
\overfullrule=0pt

\line{PhD Macroeconomics \hfil NYU}
\line{October 6, 2023 \hfil Thomas Sargent}
%
% \noindent Macroeconomics Q1 \hfill NYUAD \\
% Fall, 2019 \hfill Thomas Sargent
% \vspace{.3cm}

\centerline{\bf Practice Problem}
%\centerline{\bf Homework 3}


\medskip

\noindent{\bf Instructions:}  Please solve this fun problem.

%\noindent {\bf 1. 15 points  \quad}

\medskip
\noindent{\bf 1.  \ } 
  A consumer has preferences over consumption streams
that are ordered by
the utility functional
$$  E_0 \sum_{t=0}^\infty \beta^t u(c_t)  \leqno(1) $$
where $E_t$ is the mathematical expectation conditioned
on the consumer's time $t$ information,  $c_t$ is time $t$ consumption,
$u(c)$ is  utility function, and
$\beta \in (0,1)$ is a discount factor.  The consumer maximizes
(1) by choosing a consumption, borrowing plan
 $\{c_t, b_{t+1}\}_{t=0}^\infty$ subject to the sequence of budget constraints
$$ c_t + b_t = R^{-1} b_{t+1}  + y_t  , \quad t \geq 0$$
where $\{y_t\}$ is an exogenous
 stationary endowment process, $R$ is a constant gross
risk-free interest rate, $b_t$ is one-period risk-free  bond   maturing at
$t$, and $b_0 $ is a given initial debt that is due at time $0$.  Assume that for $t \geq 1 $, $b_t$ is chosen at time $t-1$.  Assume
that $R^{-1} = \beta$.  Assume that the endowment
process is governed by
$$ \eqalign{ y_{t+1} & = \alpha_c
+ \sum_{j=1}^2 \rho_j y_{t-j+1}  + \sum_{j=1}^2
 \delta_j  z_{t+1-j}  + \psi_1 w_{1,t+1} \cr
  z_{t+1} & = \sum_{j=1}^2 \gamma_j y_{t-j+1} + \sum_{j=1}^2 \phi_j z_{t-j+1}
+ \psi_2 w_{2,t+1}
   \cr} \leqno(1)   $$
where $\{w_{t+1}\}$ is an i.i.d.  sequence of  $2 \times 1$ vectors of random shock whose time $t+1$ component is  distributed as ${\cal N}(0,I)$
%martingale difference
%sequence,  adapted to \hfil \break $J_t = \left[\matrix{w_t & \ldots w_1 &
%c_{0} & c_{-1} & z_0 & z_{-1}\cr}\right]$, with contemporaneous
%covariance matrix $E w_{t+1}w_{t+1}' | J_t = I$, 
 and the
coefficients  $\rho_j, \delta_j, \gamma_j, \phi_j$ are such
that
the matrix
$$ A = \left[\matrix{ \rho_1 & \rho_2 & \delta_1 &\delta_2 & \alpha_c \cr
                      1 & 0 & 0 & 0 & 0 \cr
                     \gamma_1 & \gamma_2 & \phi_1 &\phi_2 & 0 \cr
                     0 & 0 & 1 & 0 & 0 \cr
                     0 & 0 & 0 & 0 & 1 \cr }\right]$$
has eigenvalues less than or equal to unity in modulus.

\medskip
\noindent {\bf a.}. Assume that the utility function $u(c) = \ln c$ %satisfies $u' > 0, u''< 0$ 
for all $c \geq 0$ and  the ``borrowing limit'' $b_t \leq \bar b$ for some $\bar b > 0$. Formulate the
consumer's problem as a dynamic program.
Write pseudo-code to solve it.
\medskip
\noindent{\bf b.}
Now  assume the
the quadratic  utility function
$$u(c_t) =  -.5 (c_t - \gamma)^2 \leqno(2) $$
where $\gamma >0$ is a bliss level of consumption.
Assume that $\bar b =+\infty$ so that there is no borrowing limit.
%
%GGHH 
%We assume that the endowment
%process has the state-space representation
%$$ \eqalign{ z_{t+1} & = \check A z_t + \check C w_{t+1}\cr
%               y_t & = \check G  z_t \cr}$$
%where $w_{t+1}$ is an i.i.d.~process with mean zero and
%identity contemporaneous covariance matrix, $\check A$ is a stable matrix,
%its eigenvalues being strictly below unity in modulus, and
%$\check G$ is a vector. The {\it state\/} vector confronting the household at $t$  is
%$\left[\matrix{b_t & z_t\cr}\right]'$, where $b_t$ is its one-period debt falling
% due at the beginning of period $t$
%and $z_t$ contains all variables useful for
%forecasting its future endowment.
Formulate the consumer's problem as a dynamic program and guess a solution.
Justify your guess.

\medskip
\noindent {\bf c.}. Continue  to assume utility function (2), but now replace the part {\bf a}   borrowing limit  with 
the following condition on the
consumption, borrowing plan:
$$  % \lim_{T \rightarrow +\infty}
 E_0 \sum_{t=0}^\infty \beta^t b_t^2 < +\infty.  $$
%This condition suffices to rule out Ponzi schemes.
Formulate the consumer's problem as a dynamic program and write pseudo-code to solve it.
  
\bye

\bye

