
\input texsis
\paper

\overfullrule=0pt
\def\frac#1#2{#1\over #2}
%
\def\bmatrix#1{\left[ \matrix{#1} \right]}
\def\be{{\beta}}
\def\toone{{t+1}}

%\def\frac#1#2{#1\over #2}
%\magnification=1200
\overfullrule=0pt

\line{Teaching tools  \hfil NYU}
\line{October 11, 2019 \hfil Thomas Sargent}

\centerline{\bf  Fun Exercises}
%\centerline{\bf Homework 3}

\medskip

\noindent{\bf Instructions:} This document consists of  ``inverse problems'' that assert answers and invite you to state  associated questions.
 Thus, for example, an  item below might be a Bellman equation for an intertemporal  optimum problem.  Your job then is  to
state completely an optimum problem associated with that Bellman equation.

\medskip
\noindent {\bf Note:} An ``intertemporal problem'' is stated in terms of  infinite sequences indexed by time $t = [0, 1, 2, \ldots, )$.




\medskip

\noindent{\bf 1.} Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 Q \cr
           \pi &  = p q - \gamma(\hat q - q)^2 } $$
so that
$$ \pi(q, Q, \hat q) = (a_0 - a_1 Q)  q - \gamma (\hat q - q)^2 .$$
Suppose that
$$ \eqalign{ \hat Q & = H(Q) \cr
            \hat q & = h(q, Q) .} $$
Please interpret the following Bellman equation in terms of a problem stated in terms of sequences  $\{p_t, q_t, Q_t\}_{t=0}^\infty$:

$$ v(q, Q) = \pi(q,Q, h(q,Q)) + \beta v\left(h(q,Q), H(Q)\right) .$$

\medskip


\noindent{\bf 2.} Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 Q \cr
           \pi &  = p q - \gamma(\hat q - q)^2 } $$
so that
$$ \pi(q, Q, \hat q) = (a_0 - a_1 Q) q - \gamma (\hat q - q)^2 .$$
Suppose that
$$ \hat Q  = H(Q)  $$
Please interpret the following Bellman equations in terms of a problem stated in terms of sequences  $\{p_t, q_t, Q_t\}_{t=0}^\infty$:

$$ v(q, Q) = \pi(q,Q, h(q, Q) ) + \beta v\left(h(q,Q), H(Q)\right) = \max_{\hat q} \{\pi(q, Q, \hat q) + \beta v(\hat q, H(Q) )   \} $$

\medskip

\noindent{\bf 3.} Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 Q \cr
           \pi &  = p q - \gamma(\hat q - q)^2 } $$
so that
$$ \pi(q, Q, \hat q) = (a_0 - a_1 Q) q- \gamma (\hat q - q)^2 .$$
Suppose that
$$ \hat Q  = H(Q)  $$
Assume that the following Bellman equations are satisfied: %in terms of problem stated in terms of sequences  $\{p_t, q_t, Q_t\}_{t=0}^\infty$:

$$ v(q, Q) = \pi(q,Q, h(q,Q)) + \beta v\left(h(q,Q), H(Q)\right) = \max_{\hat q} \{\pi(q, Q, \hat q) + \beta v(\hat q, H(Q) )   \} $$

\noindent Please interpret what it means to impose the condition

$$ h(Q, Q) = H(Q) .$$

\medskip

\noindent{\bf 4.}  There are two firms indexed by $i = 1,2$.  A subscript $i$ indicates a  firm $i$ object  and a subscript $-i$ indicates
a firm {\it not\/} $i$ object (i.e, it indicates the other firm). Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 (q_1 + q_2)  \cr
           \pi_i  &  = p q_i - \gamma(\hat q_i - q_i)^2 } $$
so that
$$ \pi(q_i, q_{-i}, \hat q_i) = [a_0 - a_1 (q_i + q_{-i}) ] q_i - \gamma (\hat q_i - q_i)^2 . $$
Let the following Bellman equations hold for $i=1,2$:
$$ v_i(q_i, q_{-i}) = \pi_i( q_i, q_{-i}, f_i(q_i, \hat q_i)) + \beta v_i(f_i(q_i, q_{-i}), f_{-i}(q_{-i}, q_i) )   $$
where
$$\eqalign{ \hat q_i & = f_i(q_i,  q_{-i}) \cr
                  \hat q_{-i} & = f_{-i}( q_{-i}, q_i)} $$
                  Please interpret the two functional equations for $ v_i(q_i, q_{-i})$ iin terms of problems  involving sequences $\{q_{1t}, q_{2t}\}_{t=0}^\infty$


\medskip

\noindent{\bf 5.}  There are two firms indexed by $i = 1,2$. ßA subscript $i$ indicates a  firm $i$ object  and a subscript $-i$ indicates
a firm {\it not\/} $i$ object (i.e, it indicates the other firm).  Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 (q_1 + q_2)  \cr
           \pi_i  &  = p q_i - \gamma(\hat q_i - q_i)^2 } $$
so that
$$ \pi(q_i, q_{-i}, \hat q_i) = [a_0 - a_1 (q_i + q_{-i}) ] q_i - \gamma (\hat q_i - q_i)^2 . $$
Let the following Bellman equations hold for $i=1,2$:
$$ v_i(q_i, q_{-i}) = \max_{\hat q_i} \{\pi_i( q_i, q_{-i}, \hat q_{i}) + \beta v_i(\hat q_i , f_{-i}(q_{-i}, q_i) )    \}$$
where
$$ \hat q_{-i} = f_{-i}( q_{-i}, q_i). $$
                  Please interpret these two Bellman equations in terms of problems  involving sequences $\{q_{1t}, q_{2t}\}_{t=0}^\infty$




\medskip
\noindent {\bf 6.}
Let $\mu_t$ be an $n \times 1$ vector, $\Sigma_t$ be an $n \times n$ positive definite matrix, and $A$ be an $n \times n$ matrix.
Let $z_1$ and $z_2$ each be $n \times 1$ random vectors. Then let  $z = \bmatrix {z_1 \cr z_2 }$ be the $2n \times 1$  multivariate normal random vector
parameterized as
$$ z \sim {\cal N}\left(\bmatrix{ \mu_t \cr A \mu_t }, \bmatrix{ \Sigma_t & \Sigma_t A' \cr
                                                                              A \Sigma_t & A \Sigma_t A' + C C' } \right) $$
By writing the distribution in this way we specify that  the mean vector of $z$ is the first argument of ${\cal N}$ and the covariance matrix of $z$ is the second argument of ${\cal N}$.
\medskip
\noindent{\bf a.}
Please find the probability distribution of $z_2$ conditional on $z_1$.
\medskip
\noindent{\bf b.} Please compute the linear least squares projection of $z_2$ on $z_1$.


\medskip
\noindent{\bf 7.} Under government policy number 1, a consumer's non-financial income $\{y_t\}_{t=0}^\infty$  is described by the state-space system
$$ \eqalign{ x_{t+1} & = A x_t + C \epsilon_{t+1} \cr
              y_t & = U_y x_t }$$
where $\epsilon_{t+1} \sim {\cal N}(0,I)$ is an i.i.d.~process.  The consumer faces a sequence of budget constraints
$$ c_t + b_t = R^{-1} b_{t+1} + y_t$$
where $R = \beta^{-1}$ is a time-invariant gross interest rate on one-period risk free debt and $b_t$ is the amount of one-period debt that
the consumer brings into period $t$, measured in time $t$ consumption goods. For each date $t \geq 0$, the consumer chooses time $t$ consumption according to  the decision rule
$$ c_t = (1-\beta) (- b_t + E_t \sum_{j=0}^\infty \beta^j y_{t+j} ) $$
where $E_t y_{t+j} = U_y A^j x_t$.  

\medskip
\noindent{\bf a.} 
 Under government policy number 2, the consumer's non-financial income process is $\{\tilde y_t\} = \{y_t\} + \{\check y_t\}$ where
 $\{y_t\}$ is the same process that prevails under government policy 1 and the $\{\check y_t\}$ process satisfies
 $$ \check y_t - \check y_{t-1} = \epsilon_t - \beta^{-1} \epsilon_{t-1} $$
 where $\epsilon_t \sim {\cal N}(0,\sigma_\epsilon^2)$ is an i.i.d.~process for $t \geq 0$  and $y_{-1}$ is a given initial condition.  
 The consumer's information set at time $t$ includes $y_{t-1}$ and $\epsilon_t, \epsilon_{t-1}, \ldots, \epsilon_0$.  
 
 \noindent{\bf b.}  Please show that the impulse response  of $y_{t+j}$ to $\epsilon_t$ is 
  $$ d_j  = \cases{ 1 &if $j = 0$;\cr
 1 -\beta^{-1}  &if $j  \geq  1$.\cr}$$


\medskip
\noindent{\bf c.} Please compute the present value of the moving average (also called `impulse response') coefficients $\{d_j\}_{j=0}^\infty$, namely,
$$ d(\beta) = \sum_{j=0}^\infty \beta^j d_j $$
where $d(z)$ is defined as the polynomial $d(z) = \sum_{j=0}^\infty  d_j z^j$.    
 
 
 
 
 \medskip
\noindent{\bf d.}  Using what you know about imposing stability conditions for  linear difference equations by ``solving unstable roots forward'', please verify the following equation:
$$ \epsilon_t = - \sum_{j=1}^\infty \beta^j (\check y_{t+j} - \check y_{t+j-1}) . $$ 
What does this formula say about the response of the present value $\sum_{j=0}^\infty\beta^j \check y_{t+j}$ to a surprise $\epsilon_t$ at time $t$?
{\it Hint:} Does it or does it not say that a surprise increase in $\check y_t$ at time $t$ is accompanied by an offsetting increase in future values of $\tilde y_{t+j},
j = 1, 2, \ldots$ that leaves the present value of income as of time $t$ unaltered?

\medskip
\noindent {\bf 8.} Consider the following model of a monopolist.  Let  $S_t$ be sales at time $t$, $Q_t$ be production output at time $t$,
 $I_t$ be inventories at the beginning of time $t$,  $\beta \in (0,1)$ be a discount factor,  $c(Q_t) = c_1 Q_t + c_2 Q_t^2$ be a cost-of-production function, where $c_1>0, c_2>0$; 
  $d(I_t, S_t) = d_1 I_t + d_2 I_t^2  + d_3 (S_t - I_t)^2$ be a cost-of-holding inventories function, where $d_1 >0, d_2 >0 $ and $d_3 >0$;
$p_t = a_0 - a_1 S_t + d_t $ be an inverse demand function for a firm's product, where $a_0>0, a_1 >0$ and $d_t$ is a demand shock at time  $t$;
   $\pi_t = p_t S_t - c(Q_t) - d(I_t, S_t) $ be the firm's profits at time $t$;
   $\sum_{t=0}^\infty \beta^t \pi_t $ be the present value of the firm's profits at time $0$;  $I_{t+1} = I_t + Q_t - S_t$ be the law of motion of inventories;
 $z_{t+1} = A_{22} z_t + C_2 \epsilon_{t+1} $ be the law of motion for an exogenous state vector $z_t$ that contains time $t$ information useful for predicting the demand shock $d_t$ where
 $\epsilon_{t+1}$ is an i.i.d.~srandom vector distributed as ${\cal N}(0,I)$; 
   $d_t = G z_t $ links the demand shock to the information set $z_t$ ;
   the constant $1$ is the first component of $z_t$.
   
   The firm wants  a production, sales, inventory-management plan $\{Q_t, S_t, I_{t+1}\}_{t=0}^\infty$ that maximizes $E_0 \sum_{t=0}^\infty \beta^t \pi_t $
   subject to $z_0,  I_0 $ being given initial conditions.
   
   
 \noindent{\bf a.}. Please formulate the firm's  problem as a discounted dynamic programming problem.  Feel free to use  the word "let" somewhere in your answer.
 
 \medskip
 
 
 \noindent{\bf b.} Please state conditions on the solution of the firm's  problem that are sufficient to imply that as $t \rightarrow + \infty$ $\bmatrix{z_t \cr I_t} $ converges  to a normally distributed random vector $x_\infty \sim {\cal N}(\mu, \Omega)$.
 
 \medskip
\noindent{\bf c.} Please give formulas for $\mu$ and $\Omega $ in terms of the firm's objective and its constraints.
   
   \newpage
   
\medskip
\noindent{\bf 9.} Let $C(Q) = c_1 Q + c_2 Q^2$ where $c_1, c_2 > 0$; $d(I, S) = d_1 I + d_2 I^2 + d_3 (S - I)^2$ where $d_1>0, d_2 >0, d_3 >0$; 
$a_1 >0, a_2 >0$ and $\pi(S, Q, I, z) = (a_0 - a_1 S + G z) S - c(Q) - d(I,S)$.  Suppose that $\phi(\epsilon)$ is a probability density for an $m$-dimensional vector
$\epsilon$ with mean vector $0$ and covariance matrix $I$. Consider the following functional  equation in $v(I, z)$:
$$ v(I, z) = \max_{Q, S} = \left\{ \pi(S, Q, I, z) + \beta \int v(I+Q - S, Az + C \epsilon) \phi(\epsilon) d \epsilon \right\} . $$
Please describe a problem in a space of sequences of $\{S_t, Q_t, I_t, z_t\}_{t=0}^\infty$ for which    the above functional equation provides an answer.

\medskip
\noindent{\bf 10.} This is an exercise about the meaning of ``i.i.d.''
\medskip
\noindent{\bf a.} Define a model.

\medskip

\noindent Time is discrete and takes values $t =0, 1, \ldots$.  Let $F = \{f_j\}_{j=0}^N$ be  a discrete probability distribution that satisfies $f_j > 0$ for $j=0, \ldots, N$, $\sum_{j=0}^N f_j = 1$,
where $f_j = {\rm Prob}(W = j)$.  Let $G = \{g_j\}_{j=0}^N$ be  another  discrete probability distribution that satisfies $g_j > 0$ for $j=0, \ldots, N$, $\sum_{j=0}^N g_j = 1$,
where $g_j = {\rm Prob}(W = j)$.  A state variable $s_t \in \{0,1\}$, meaning that $s_t$ takes value either $0$ or $1$.  The state variable follows the law of motion
$$ s_{t+1} = s_t \quad \forall t \geq 0 .$$
Also, the initial state $s_0$ is a random variable with ${\rm Prob} (s_0 = 1) = \pi \in (0,1)$.  

\medskip
\noindent{\bf b.}  Let  $\{w_t\}_{t=0}^\infty$ be a sequence formed as independent draws from the distribution $G$, meaning that 
for any sequence of $i_0, i_1, i_2, \ldots $ of integers $\{i_t\}$ where the $t$th element $i_t  \in \{0, 1, \ldots, N\}$, 
 $${\rm Prob}(w_0 = i_0, w_2 = i_1, w_2, \ldots)  = g_{i_0} g_{i_1} g_{i_2}  \cdots $$
(i.e., successive draws are from the same distribution and don't depend on earier draws). Please compute the mean of $w_t$ for $t \geq 0$.  
For $t \geq 1$, 
please compute the probability distribution of  $w_t$ conditional on the history $w^{t-1} = w_{t-1}, w_{t-2}, \ldots, w_0$. 

\medskip
\noindent Now let  the probability distribution of $w_t$  be given by
$$ h_j = {\rm Prob}(w_t = j) = s_t f_j + (1-s_t) g_j ,  \quad \forall j \geq 0. $$
Assume that you know the three probability distibutions $G, F, \pi$.
Please answer these questions about the associated   model of  $\{w_t\}_{t=0}^\infty$.



\medskip
\noindent{\bf c.} For $t \geq 1$, 
please compute the probability distribution of  $w_t$ conditional on the history $w^{t-1} = w_{t-1}, w_{t-2}, \ldots, w_0$. 

\medskip 
\noindent{\bf d.} Under the model implied by $G, F, \pi$, is $\{w_t\}_{t=0}^\infty$ an independently and identically distributed sequence of random variables?


\noindent The following questions change conditioning information in an important way; depending on how you define ``model'', perhaps they also change the ``model''.  


\medskip
\noindent{\bf e.} For $t \geq 1$, 
please compute the probability distribution of  $w_t$ conditional on  the history $w^{t-1} = w_{t-1}, w_{t-2}, \ldots, w_0$ AND $s_0$. 

\medskip 
\noindent{\bf f.} Under the ``model'' implied by $G, F, s_0$, is $\{w_t\}_{t=0}^\infty$ an independently and identically distributed sequence of random variables?




 












\bye



\medskip

\noindent{\bf 1.}

\medskip
Assume that  $F(P)$ is a right continuous function mapping $[0, B]$ into $[0,1]$ and that the real scalar $c > 0$.  For what intertemporal problem is the following
a Bellman equation?
$$ v(p) = \min \{p, c + \int v(p') d F(p') \}$$.  A state variable $s_t \in \{0,1\}$ takes one of two values, namely, $0$ or $1$.  A probability


\medskip



\noindent {\bf 2.}  Assume that  $F$ is a right continuous function mapping $[0, B]$ into $[0,1]$ and that $\beta \in (0,1)$. For what intertemporal problem is the following a Bellman equation?
$$ v(w) = \max\left\{ {w\over{1-\beta}},  \beta \int v(w') d F^2(w') \right\} $$



\medskip
\noindent{\bf 3.} Where $x \in {\bf R}$, $f >0, d >0 $,   $\beta \in (0,1)$, and $g \in  {\bf R} \rightarrow {\bf R}$, for what intertemporal problem is the following  a Bellman equation?
$$ v(x) =  \{- f x^2 - d(gx  - x)^2 + \beta v(gx) \} $$


\medskip
\noindent{\bf 4.}   Where $x \in {\bf R}$, $f >0, d >0 $, and   $\beta \in (0,1)$, for what intertemporal problem is the following  a Bellman equation?
$$ v(x) = \max_{x'} \{ - f x - d(x' - x)^2 + \beta v(x') \} $$


\medskip
\noindent {\bf 5.} Where $x \in {\bf R}$, $f >0, d >0 $, $\sigma >0$,    $\beta \in (0,1)$, and $\phi$ is a univariate normal density function for a random
variable with mean $0$ and variance $1$, for what intertemporal problem is the following  a Bellman equation?
$$ v(x) = \max_{u} \{ - f x - d u^2 + \beta \int v(x +u + \sigma \epsilon) d \phi(\epsilon) \} $$



\medskip



\noindent {\bf 6.}  Assume that  $F$ and $G$ are distinct  right continuous functions each mapping $[0, B]$ into $[0,1]$, that $\pi \in (0,1)$, and  that $\beta \in (0,1)$. For what intertemporal problem is the following a Bellman equation?
$$ v(w) = \max\left\{ {w\over{1-\beta}},  \beta [  \pi \int v(w') d F(w') + (1-\pi) \int v(w') d G(w') ]  \right\} $$


\bye
