
\input texsis
\paper

\overfullrule=0pt
\def\frac#1#2{#1\over #2}
%
\def\bmatrix#1{\left[ \matrix{#1} \right]}
\def\be{{\beta}}
\def\toone{{t+1}}

%\def\frac#1#2{#1\over #2}
%\magnification=1200
\overfullrule=0pt

\line{Econ 311 \hfil U of Chicago}
\line{February 4, 2021 \hfil Lars and Tom}

\centerline{\bf  Fun Problems}


\medskip

\noindent{\bf -1.} Let $x_{t+1}, a_{t+1}, y_{t+1}$ all be scalar stochastic processes obeying the 
state-space representation
$$\eqalign{ x_{t+1} & = a x_t + b a_{t+1} \cr
            y_{t+1} & = d x_t + g a_{t+1} }$$
where $x_t$ is a hidden Markov state, $a_t \sim {\cal N}(0, \sigma_a^2)$is the time $t$ component
of an i.i.d. sequence of random shocks, $x_0 \sim {\cal N}(0, \sigma_0)$
 and at time $t$ a history $y^t =[] y_t, y_{t-1}, \ldots, ]$ is observed.  

\medskip
\noindent{\bf a.} Please find a time-invariant {\it innovations representation\/} corresponding to the preceding state
space representation.  

\medskip
\noindent {\bf b.} Using the innovation representation,  please compute the conditional
expectation $E [x_{t+1} | y^{t+1}]$. 

\medskip


\noindent{\bf 0.} Consider the follow special case of a classic Brock-Mirman growth model that 
we assume describes economic history of a particular country called A.
Let $k_t$ be a stock of physical capital, $c_t$ be a consumption rate, and the transition law for the capital 
stock 
$$ k_{t+1} + c_t = A k_t^\alpha \theta_t, \quad \alpha \in (0,1) $$
where $k_0$ is an initial condition and $\ln (\theta_t) \sim {\cal N}(0, \sigma_\theta^2)$ is the time $t$
component of an i.i.d. sequence of technology shocks.  Assume that consumption is described by
the decision rule
$$ c_t = (1-\beta \alpha) A k_t^\alpha \theta_t, \quad \beta \in (0,1) $$
for all $t \geq 0$.

An economist at the World Bank knows the parameters $A, \beta, \alpha, \sigma_\theta$ and the structure of
the model.  The economist does not know the initial capital stock $k_0$ or subsequent capital stocks
$\{k_{t+1}\}_{t=0}^T$ for some large $T$.  However, for country A the World Bank economist has an historical record
of observations on $\log \tilde c_t, t=0, \ldots, T$, where 
$$ \tilde c_t = c_t \epsilon_t $$
where $\log \epsilon_t \sim {\cal N}(0, \sigma_\epsilon^2)$ is the time $t$ component of an i.i.d. sequence of
of measurement errors.  The stochastic processes $\{\theta_t\}_{t=0}^T$ and $\{\epsilon_t\}_{t=0}^T$ are mutually
independent at all dates.  The economist knows the measurement error parameter $\sigma_\epsilon$ along with the 
other parameters already mentioned.  
The economist's job is to  estimate  the capital stock process $\{k_t\}_{t=0}^T$ for country $A$. 

 The economist expresses that he/she does not know $k_0$ by instead  pretending that 
he/she knows very  little. In particular, he/she acts as if he/she has a prior probability distribution
 $\log k_0 \sim {\cal N}(0, 10000)$. The prior variance  $10000$ is so very large that the implied
distribution of $\log k_0$ covers all capital stocks that World Bank economists have ever reported ``seeing''.  

\medskip
\noindent{\bf a.} The economist's supervisor asks him/her to justify his/her choice of prior for $\log k_0$.
Being a person of few words, the economist says: ``conjugate prior''. Does this make sense? Please elaborate
on the economist's justification.

\medskip
\noindent{\bf b.} Please describe a recursive algorithm for computing the conditional mathematical
expectation $E \log k_t | h_t$ where $h_t = []\log \tilde c_t, \log \tilde c_{t-1}, \ldots, \log \tilde c_0]$.

\medskip
\noindent {\bf c.} Please find a Hidden Markov Model for the $\{\log k_t\}_{t=0}^T$ process.





\medskip

\noindent{\bf 1.}  \quad Consider the
modified version of the optimal linear regulator problem where the
objective is to maximize
$$-  \sum_{t=0}^\infty  \left\{ x_t' R x_t + u_t' Q u_t
    + 2 u_t' H x_t \right\}$$
subject to the law of motion:
$$ x_{t+1} = A x_t + B u_t .$$
Here $x_t$ is an $n \times 1$ state vector, $u_t$ is a $k \times 1$
vector of controls, and $x_0$ is a given initial condition.   The matrices
$R, Q$ are positive definite and symmetric.
 The
maximization is with respect to sequences $\{u_t, x_t\}_{t=0}^\infty$.
\medskip
\noindent {\bf a.} Show  that the optimal policy has the form
$$ u_t = - (Q +  B'P B)^{-1} (B' P A + H) x_t,$$
where $P$ solves the
algebraic matrix Riccati equation
$$ P = R +  A' P A - ( A' P B + H')
       (Q+ B'PB)^{-1} ( B' P A + H) . \eqno(1) $$ %\EQN ricccross $$
\medskip
\noindent{\bf b.}  You have to solve such a control problem fast but have no computer resources available personally.
But you do have a research assistant from the statistics department who knows nothing about linear quadratic optimal 
control theory. But she is an expert on Kalman filtering and has fast programs ready to go.  Can you give
her instructions that will trick  her programs into solving your problem?



\medskip
\noindent {\bf 2.}
Let $\mu_t$ be an $n \times 1$ vector, $\Sigma_t$ be an $n \times n$ positive definite matrix,  
$A$ be an $n \times n$ matrix, and $B$ be an $n \times k$ matrix.
Let $z_1$ and $z_2$ each be $n \times 1$ random vectors. Then let  $z = \bmatrix {z_1 \cr z_2 }$ be the $2n \times 1$
  multivariate normal random vector
parameterized as
$$ \bmatrix {z_1 \cr z_2 } \sim {\cal N}\left(\bmatrix{ \mu_t \cr A \mu_t }, \bmatrix{ \Sigma_t & \Sigma_t A' \cr
                                                                              A \Sigma_t & A \Sigma_t A' + B B' } \right) $$
%By writing the distribution in this way we specify that  the mean vector of $z$ is the first argument of ${\cal N}$ and the covariance matrix of $z$ is the second argument of ${\cal N}$.
\medskip
\noindent{\bf a.}
Please find the probability distribution of $z_2$ conditional on $z_1$.
\medskip
\noindent{\bf b.} Please compute the linear least squares projection of $z_2$ on $z_1$.


\medskip

\noindent{\bf 3.} Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 Q \cr
           \pi &  = p q - \gamma(\hat q - q)^2 } $$
so that
$$ \pi(q, Q, \hat q) = (a_0 - a_1 Q)  q - \gamma (\hat q - q)^2 .$$
Suppose that
$$ \eqalign{ \hat Q & = H(Q) \cr
            \hat q & = h(q, Q) .} $$
Please interpret the following Bellman equation in terms of a problem stated in terms of sequences  $\{p_t, q_t, Q_t\}_{t=0}^\infty$:

$$ v(q, Q) = \pi(q,Q, h(q,Q)) + \beta v\left(h(q,Q), H(Q)\right) .$$

\medskip


\noindent{\bf 4.} Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 Q \cr
           \pi &  = p q - \gamma(\hat q - q)^2 } $$
so that
$$ \pi(q, Q, \hat q) = (a_0 - a_1 Q) q - \gamma (\hat q - q)^2 .$$
Suppose that
$$ \hat Q  = H(Q)  $$
Please interpret the following Bellman equations in terms of a problem stated in terms of sequences  $\{p_t, q_t, Q_t\}_{t=0}^\infty$:

$$ v(q, Q) = \pi(q,Q, h(q, Q) ) + \beta v\left(h(q,Q), H(Q)\right) = \max_{\hat q} \{\pi(q, Q, \hat q) + \beta v(\hat q, H(Q) )   \} $$

\medskip

\noindent{\bf 5.} Where $a_0 >0, a_1> 0, \gamma > 0$ and $\beta \in (0,1)$ suppose that
$$\eqalign{ p & = a_0 - a_1 Q \cr
           \pi &  = p q - \gamma(\hat q - q)^2 } $$
so that
$$ \pi(q, Q, \hat q) = (a_0 - a_1 Q) q- \gamma (\hat q - q)^2 .$$
Suppose that
$$ \hat Q  = H(Q)  $$
Assume that the following Bellman equations are satisfied: %in terms of problem stated in terms of sequences  $\{p_t, q_t, Q_t\}_{t=0}^\infty$:

$$ v(q, Q) = \pi(q,Q, h(q,Q)) + \beta v\left(h(q,Q), H(Q)\right) = \max_{\hat q} \{\pi(q, Q, \hat q) + \beta v(\hat q, H(Q) )   \} $$

\noindent Please interpret what it means to impose the condition

$$ h(Q, Q) = H(Q) .$$



\bye

